Formalizing Stack Safety as a Security Property

### 2023-01-31 submission5/core.tex
Moved things around in the introduction and went back to v1 of key ideas. Also trimmed down v1 even further.
Moved advanced machine later again.

\input{ideas-v2}

@@

The definition of \(Op\) is most convenient to present decomposed into
rules for each operation. We have already seen the intuition behind the rules for
\(\mathbf{alloc}\), \(\mathbf{call}\), and \(\mathbf{ret}\).
The full machine defines additional operations, and extends some of the old ones.
We add a field \(\overline{sa}\) to the \(\mathbf{call}\) operation: a set of
triples of a register \(\reg\), a base offset \(\mathit{off}\), and a size \(sz\), denoting
that the object at the offset \(\mathit{off}\) from \(\reg\) is to be passed as an argument.
The new \(\mathbf{tailcall}\) operation has the same fields, but deals with tail-call
optimizations in which the callee will reuse the caller's stack. The \(\mathbf{alloc}\) operation
gains a boolen flag indicating whether the allocation is to be public, accessible to
all functions until deallocated. The rules are given in \cref{fig:advops}; the
example is equivalent to these rules with \(\mathbf{call}\) always instantiated with
\(\overline{sa}\) as the empty set, and \(\mathbf{alloc}\) with \(\mathbf{f}\).

@@

\apt{Remainder of this section should be moved to AFTER the rest of the formalization.}

\subsection{Sharing Stack Memory}
In our examples so far, we have presented a vision of stack safety in which
the interface between caller and callee is in the registers that pass
arguments and return values. This is frequently not the case in a realistic
setting. Arguments may be passed on the stack due to being spilled, as an implementation
of variadic arguments, or because they are reference objects that inherently have
pass-by-reference semantics. The caller may also take the address of a local object and
pass that as a pointer to its callee.

@@

\apt{Duplicates title of last section.}

@@

>>>>

\input{ideas-v1}

@@

The definition of \(Op\) is most convenient to present decomposed into
rules for each operation. We have already seen the intuition behind the rules for
\(\mathbf{alloc}\), \(\mathbf{call}\), and \(\mathbf{ret}\).
For the machine described in the example, the \(Op\) rules would be those
found in \cref{fig:basicops}.

\begin{figure}
  \judgmentbr{\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap
    \{\addr \mid V ~ \addr = \unsealed\}\)}
             {\(V' = V \llbracket \addr \mapsto \sealed \mid \addr \in \components \rrbracket\)}
             {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{f} ~ \mathit{off, sz}) ~ (V,\sigma) = (V',\sigma)\)}
    %
    \judgmentbr{\(V' = V \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket
                 \llbracket \reg \mapsto \public | \reg \in \overline{\reg_{args}} \rrbracket\)}
               {\(V'' = V'\llbracket \addr \mapsto \sealed | V' ~ \addr = \object \land \addr \not \in (\mathit{passed} ~ \overline{sa} ~ \mach) \rrbracket\)}
               {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
                 ~ (V,\sigma) = (V'',V::\sigma)\)}
    %
    \judgment{\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
             {\(Op ~ \mach ~ \mathbf{return} ~ (\_, \sigma) = (V, \sigma)\)}
  \caption{Basic Operations}
  \label{fig:basicops}
\end{figure}

@@

@@

@@

\section{Expanding the Machine}

The system we've modeled so far has been very simple, but our properties
support an important dimension of generality: adding features. We will now
briefly introduce the features of our primary testing machine.

The full machine defines additional operations, and extends some of the old ones.
We add a field \(\overline{sa}\) to the \(\mathbf{call}\) operation: a set of
triples of a register \(\reg\), a base offset \(\mathit{off}\), and a size \(sz\), denoting
that the object at the offset \(\mathit{off}\) from \(\reg\) is to be passed as an argument.
The new \(\mathbf{tailcall}\) operation has the same fields, but deals with tail-call
optimizations in which the callee will reuse the caller's stack. The \(\mathbf{alloc}\) operation
gains a boolen flag indicating whether the allocation is to be public, accessible to
all functions until deallocated. The rules are given in \cref{fig:advops}; the
rules in \cref{fig:basicops} can be recaptured by instantiating
\(\mathbf{call}\) with \(\overline{sa}\) as the empty set, and \(\mathbf{alloc}\)
with \(\mathbf{f}\).

\subsection{Sharing Stack Memory}
In our examples, we have presented a vision of stack safety in which
the interface between caller and callee is in the registers that pass
arguments and return values. This is frequently not the case in a realistic
setting. Arguments may be passed on the stack due to being spilled, as an implementation
of variadic arguments, or because they are reference objects that inherently have
pass-by-reference semantics. The caller may also take the address of a local object and
pass that as a pointer to its callee.
